Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm

### 2022-04-21 final_metrics_analysis.Rmd
making faceted priors plot

# for Dianne
use_condaenv("base")

@@

```{python crosstable of recitivism by race}
crosstable=pd.crosstab(df_viz['race'],df_viz['two_year_recid'])
crosstable
```

```{python exploratory recidivism mosaic plot}
props={}
props[('Caucasian','Recidivated')]={'facecolor':'green', 'edgecolor':'white'}
props[('Caucasian','Survived')]={'facecolor':'green', 'edgecolor':'white'}
props[('African-American','Recidivated')]={'facecolor':'xkcd:purple','edgecolor':'white'}
props[('African-American','Survived')]=        {'facecolor':'xkcd:purple','edgecolor':'white'}
labelizer=lambda k:{('Caucasian','Recidivated'):822,('African-American','Recidivated'):1661,('Caucasian','Survived'):1278,('African-American','Survived'):1512}[k]
mosaic(df_viz,['race','two_year_recid'],labelizer=labelizer,properties=props)
plt.show()
```

@@

```{python wordcloud setup}
from os import path
from PIL import Image
from wordcloud import WordCloud, STOPWORDS, ImageColorGenerator
```

```{python wordcloud}
text = " ".join(charge for charge in df_viz.c_charge_desc)
stopwords = set(STOPWORDS)
stopwords.update(["arrest", "case"])
wordcloud = WordCloud(stopwords=stopwords, background_color="white").generate(text)
```

```{python wordcloud image}
plt.imshow(wordcloud, interpolation='bilinear')
plt.axis("off")
plt.show()
```

@@

```{python exploratory recidivism by race table}
# table of recidivism by race
pd.crosstab(index = df_viz["race"], columns = df_viz["two_year_recid"])
```

```{python exploratory priors by race}
priors_by_race = sns.countplot(x="race", color="priors_count", data=df_viz, palette=workshop_palette)

plt.title('Prior Charges Count by Race')
plt.show()
```

```{python exploratory priors by sex}
priors_by_race = sns.countplot(x="sex", color="priors_count", data=df_viz, palette=workshop_palette)

plt.title('Prior Charges Count by Sex')
plt.show()
```

>>>>

# for Dianne
# use_condaenv("base")

@@

```{python crosstable of recidivism by race}
crosstable=pd.crosstab(df_viz['race'],df_viz['two_year_recid'])
crosstable
```

```{python exploratory recidivism mosaic plot}
props={}
props[('Caucasian','Recidivated')]={'facecolor':'red', 'edgecolor':'white'}
props[('Caucasian','Survived')]={'facecolor':'red', 'edgecolor':'white'}
props[('African-American','Recidivated')]={'facecolor':'xkcd:aqua','edgecolor':'white'}
props[('African-American','Survived')]=        {'facecolor':'xkcd:aqua','edgecolor':'white'}
labelizer=lambda k:{('Caucasian','Recidivated'):822,('African-American','Recidivated'):1661,('Caucasian','Survived'):1278,('African-American','Survived'):1512}[k]
mosaic(df_viz,['race','two_year_recid'],labelizer=labelizer,properties=props)
plt.show()
```

@@

@@

```{python exploratory priors by race}
priors_by_race = sns.displot(x="priors_count", row = "race", col = "sex", hue ="race", data=df_viz, palette=workshop_palette)

#plt.title("Prior Charges Count by Race and Sex)
plt.show()
```
